%%%%%%%% ICML 2026 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2026} with \usepackage[nohyperref]{icml2026} above.
\usepackage{ninecolors}
\usepackage[colorlinks=true,
            linkcolor=blue4,
            filecolor=blue4,
            urlcolor=blue4,
            citecolor=blue4]{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2026}

% For preprint, use
% \usepackage[preprint]{icml2026}

% If accepted, instead use the following line for the camera-ready submission:
% \usepackage[accepted]{icml2026}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{thmtools}
\usepackage{thm-restate}

% special macros
\input{amacros.tex}
\input{dmacros.tex}

% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \theoremstyle{plain}
% \newtheorem{theorem}{Theorem}[section]
% \newtheorem{proposition}[theorem]{Proposition}
% \newtheorem{Lem}{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \theoremstyle{definition}
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{assumption}[theorem]{Assumption}
% \theoremstyle{remark}
% \newtheorem{remark}[theorem]{Remark}

\theoremstyle{plain}
\declaretheorem[numberwithin=section,name=Theorem]{theorem}
\declaretheorem[sibling=theorem,name=Proposition]{proposition}
\declaretheorem[sibling=theorem,name=Lemma]{lemma}
\declaretheorem[sibling=theorem,name=Corollary]{corollary}
\theoremstyle{definition}
\declaretheorem[sibling=theorem,name=Definition]{definition}
\declaretheorem[sibling=theorem,name=Assumption]{assumption}
\theoremstyle{remark}
\declaretheorem[sibling=theorem,name=Remark]{remark}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}


% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Linear contextual bandits with paid observations}

\begin{document}

\twocolumn[
  \icmltitle{Best-of-Both Worlds for linear contextual bandits with paid observations}

  % It is OKAY to include author information, even for blind submissions: the
  % style file will automatically remove it for you unless you've provided
  % the [accepted] option to the icml2026 package.

  % List of affiliations: The first argument should be a (short) identifier you
  % will use later to specify author affiliations Academic affiliations
  % should list Department, University, City, Region, Country Industry
  % affiliations should list Company, City, Region, Country

  % You can specify symbols, otherwise they are numbered in order. Ideally, you
  % should not use this facility. Affiliations will be numbered in order of
  % appearance and this is the preferred way.
  \icmlsetsymbol{equal}{*}

  \begin{icmlauthorlist}
    \icmlauthor{Antoine Moulin}{upf}
    \icmlauthor{Nathan Boyer}{oxford}
    \icmlauthor{Patrick Rebeschini}{oxford}
    \icmlauthor{Dorian Baudry}{inria}
  \end{icmlauthorlist}

  \icmlaffiliation{upf}{Universitat Pompeu Fabra, Barcelona, Spain}
  \icmlaffiliation{oxford}{Department of Statistics, University of Oxford, Oxford, UK}
  \icmlaffiliation{inria}{Univ.\ Grenoble Alpes, Inria, CNRS, Grenoble INP, LIG, 38000 Grenoble, France}

  \icmlcorrespondingauthor{Dorian Baudry}{dorian.baudry@inria.fr}

  % You may provide any keywords that you find helpful for describing your
  % paper; these are used to populate the "keywords" metadata in the PDF but
  % will not be shown in the document
  \icmlkeywords{Linear contextual bandits, best-of-both-worlds, paid observations}

  \vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column listing the
% affiliations and the copyright notice. The command takes one argument, which
% is text to display at the start of the footnote. The \icmlEqualContribution
% command is standard text for equal contribution. Remove it (just {}) if you
% do not need this facility.

% Use ONE of the following lines. DO NOT remove the command.
% If you have no special notice, KEEP empty braces:
\printAffiliationsAndNotice{}  % no special notice (required even if empty)
% Or, if applicable, use the standard equal contribution text:
% \printAffiliationsAndNotice{\icmlEqualContribution}

\begin{abstract}
  We study the problem of linear contextual bandits with paid observations, where at each round the learner selects an action in order to minimize its loss in a given context, and can then decide to pay a fixed cost to observe the loss of any arm. Building on the Follow-the-Regularized-Leader framework with efficient estimators via Matrix Geometric Resampling, we introduce a computationally efficient Best-of-Both-Worlds (BOBW) algorithm for this problem. We show that it achieves the minimax-optimal regret of $\Theta(T^{2/3})$ in adversarial settings, while guaranteeing poly-logarithmic regret in (corrupted) stochastic regimes. Our approach builds on the framework from \cite{BOBWhardproblems} to design BOBW algorithms for ``hard problem'', using analysis techniques tailored for the setting that we consider. %, which may be of independent interest.
\end{abstract}

\input{sections/introduction.tex}
\input{sections/preliminaries.tex}
\input{sections/algorithm.tex}
\input{sections/analysis.tex}
\input{sections/conclusion.tex}

% Acknowledgements should only appear in the accepted version.
% \section*{Acknowledgements}

% \textbf{Do not} include acknowledgements in the initial version of the paper
% submitted for blind review.

% If a paper is accepted, the final camera-ready version can (and usually should)
% include acknowledgements.  Such acknowledgements should be placed at the end of
% the section, in an unnumbered section that does not count towards the paper
% page limit. Typically, this will include thanks to reviewers who gave useful
% comments, to colleagues who contributed to the ideas, and to funding agencies
% and corporate sponsors that provided financial support.

\section*{Impact Statement}

This paper presents work whose goal is to advance the field of Machine
Learning. There are many potential societal consequences of our work, none
which we feel must be specifically highlighted here.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite


\bibliography{biblio}
\bibliographystyle{icml2026}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn


\section{Comments Dorian}
\DB{Updated!}

Comments after visit in Oxford, Feb 26:
\begin{itemize}
\item We cleaned the beginning of the analysis (FTRL stability/penalty/bias decomposition) with the ghost sample trick.
\item We explored new ways to tune the FTRL approach in order to fit either in the setting of Tsuchiya or Kuroki
\item We have good hope to complete a BoBW analysis in both settings, but technical details still need to be sorted out. 
\end{itemize}

\paragraph{Direction 1: fitting Tsuchiya's framework} The main problem here is the stability term: the covariance matrix essentially involves the observation probabilities on all contexts, while we would like to cancel the probability on the ghost sample $X_0$. To solve this, we proposed a \textit{fixed} observation probabilities for all contexts, that cancels the \textit{average} sampling probability. This is computationnally burdensome, but fits quite elegantly in Taira's framework from a theoretical standpoint. Perhaps surprisingly, it seems that we can get the $\log(T)$ stochastic bound with this approach, but with rather ugly constants.  

\paragraph{Direction 2: fitting Kuroki's framework} After processing the SPB matching analysis it appeared even more surprising to us that Kuroki's approach doesn't get the problem of ghost sample/current context mismatch. Hence, Wwe investigates a learning rate tuning similar to theirs to see if we would allow results that resemble theirs. It appears that we can, but we still need to process the technical details (again, mostly for the self-bounding/stochastic aspect). This definitely looks promising.

Overall, I think that if we can close the analysis we could develop the two approaches in the paper and present them as kinda complementary: clean SPB matching/better asymptotic stochastic bound but more heavy, versus efficient update at the cost of less clean stochastic bound.

It remains to understand if the second approach is really fundamentally different to the first one. Without contexts we have the intuition that Kuroki-style tuning kinda approximates Tsuchiya clean SPB matching. Essentially, they are matching the targeted \textit{upper bounds} of the SPB terms instead of the terms themselves directly. Having experiments would definitely be valuable to understand the practical benefits and drawbacks of each approach. 

\paragraph{To-DO} Antoine will take the technical lead before Neurips deadline, and we'll try together to make the two directions mentioned above work. Main tasks are:
\begin{itemize}
	\item Fill the gaps in the proof. There are various placeholders in the draft now. 
	\item Seems rather clear that both direction can yield the adversarial bound, but the stochastic ones are still very sketched in the draft. Process the details, this is the main challenge.  	
\end{itemize}

\paragraph{Some Questions}
\begin{itemize}
	\item It is not clear that we can relax the assumption that $\Delta_{\min}>0$ (even to "just" get the $T^{2/3}$). Doing so even in Yuko's setting (without costs) might be interesting. I have a feeling that the entropies used in the literature are inherently unable to do this because they don't decrease when two arms have constant probability, even if the others are converging to zero. Or maybe this can be solved with arm-dependent learning rate (this seems to be Haipeng's solution). It would be interesting to further refine what it can mean to be BOBW in a continuous contexts setting, and might be cool to check the papers on Greedy to see if there share nice insights for the stochastic case.
	\item Not clear what the lower bound should be (even in the full-info/no cost case), might be interesting too. 
\end{itemize}

\begin{itemize}
    \item Kuroki et al. have a worst stochastic bound in $(\log(T))^3+d(\log(T))^2$. Seems to be because how the exploration parameter should both be large enough for the MGR bias and at the same time depend tightly on the learning rate.
\end{itemize}

\input{sections/omitted_proofs.tex}
\clearpage
\input{sections/technical_tools.tex}
\clearpage
\input{sections/mgr.tex}
\clearpage
\input{sections/complexity.tex}
\clearpage
\input{sections/draft_nathan.tex}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
% Modified again in 2023 and 2024 by Sivan Sabato and Jonathan Scarlett.
% Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.
