\section{Tentative regret analysis/algorithm design}

    \subsection{Notation}

\antoine{Writing new notation here, propagate back into the main text later} For $\alpha \in \spr*{0, 1}$, we define the negative Tsallis entropy with parameter $\alpha$ over the nonnegative orthant $\bbRnnK$ as
%
\begin{equation*}
    H_\alpha \spr*{q} = \frac{1}{1 - \alpha} \sumaK \spr*{q_a - q_a^\alpha}.
\end{equation*}
%
\db{Mettre $1/\alpha$}
Note that $\nabla H_\alpha \spr*{q} = \frac{1}{1 - \alpha} \spr*{\bfone - \alpha q^{\alpha - 1}}$, and $\nabla^2 H_\alpha \spr*{x} = \alpha \Diag \spr*{q^{\alpha - 2}}$. At time $t$, for any context $x$ and arm $a$, the losses are defined as $\ell_t \spr*{x, a} = \inp*{x, \theta_{t, a}}$. We denote $p_t \spr*{x}$ the probability of observing each arm at time $t$ for context $x$, \ie, $p_t \spr*{x} = \bbP_t \sbr*{a \in O_t \given X_t = x}$, and we let $\Sigma_t = \bbE_t \sbr*{p_t \spr*{X_t} X_t X_t\transpose}$ the (conditional) covariance matrix of the contexts for the observed actions. We denote $\ptmin = \min_{x \in \cX} p_t \spr*{x}$. We consider the corresponding importance-weighted estimator $\wh{\theta}_{t, a} = \Sigma_t^{-1} X_t \ell_t \spr*{X_t, a} \, \ind_{\scbr*{a \in O_t}}$ and note that this is an unbiased estimator
%
\begin{align*}
    \bbE_t \sbr*{\wh{\theta}_{t, a}} &= \Sigma_t^{-1} \bbE_t \sbr*{X_t X_t\transpose \theta_{t, a} \ind_{\scbr*{a \in O_t}}}\\
    &= \Sigma_t^{-1} \bbE_t \sbr*{X_t X_t\transpose \theta_{t, a} \bbE_t \sbr*{\ind_{\scbr*{a \in O_t}} \given X_t}}\\
    &= \Sigma_t^{-1} \bbE_t \sbr*{p_t \spr*{X_t} X_t X_t\transpose} \theta_{t, a}\\
    &= \theta_{t, a},
\end{align*}
%
where we used the definition of the losses $\ell_t$, the law of total expectation, and the definitions of $p_t$ and $\Sigma_t$. We further denote $\wt{\theta}_{t, a}$ the MGR approximation of $\wh{\theta}_{t, a}$, defined as
%
\begin{equation*}
    \wt{\theta}_{t, a} = \whsigmaplust X_t \ell_t \spr*{X_t, a} \, \ind_{\scbr*{a \in O_t}},
\end{equation*}
%
where $\whsigmaplust$ is the random matrix output by the MGR procedure in Algorithm~\ref{alg:MGR}. This is a biased estimator, and assuming the procedure sampled $M_t$ additional contexts, we have \antoine{check the last equation carefully}
%
\begin{align*}
    \bbE_t \sbr*{\wt{\theta}_{t, a}} &= \bbE_t \sbr*{\whsigmaplust} \bbE_t \sbr*{X_t \ell_t \spr*{X_t, a} \, \ind_{\scbr*{a \in O_t}}}\\
    &= \bbE_t \sbr*{\whsigmaplust} \Sigma_t \cdot \theta_{t, a}\\
    &= \theta_{t, a} - \spr*{I - \frac12 \Sigma_t}^{M_t} \theta_{t, a}.
\end{align*}
%
We denote $\wh{\ell}_t$ and $\wt{\ell}_t$ the corresponding losses. For any $x$, any $t$, we consider the policies $q_t$ defined as
%
\begin{equation*}
    q_t \spr*{\cdot \given x} = \argmin_{q \in \Delta_K} \scbr*{\sum_{s=1}^{t-1} \inp*{q, \wt \ell_s \spr*{x}} + \psi_t \spr*{q} + \bar\psi \spr*{q}},
\end{equation*}
%
where we defined $\psi_t \spr*{q} = \frac{1}{\eta_t} H_\alpha \spr*{q}$ for some $\eta_t > 0$ to be defined, $\bar\psi \spr*{q} = \bar\beta H_{\bar\alpha} \spr*{q}$ with $\bar\alpha = 1 - \alpha$ and some $\bar\beta > 0$ to be defined, and $\wt \ell_s \spr*{x} \in \bbR^K$ is the vector of estimated losses at time $s$ for context $x$, with components $\wt \ell_s \spr*{x, a} = \inp*{x, \wt \theta_{s, a}}$ for any $a \in \sbr*{K}$.



\subsection{Analysis}

\DB{A suggestion: maybe first do a clean appendix with the regret decomposition (up to the FTRL lemma) with the ghost sample technique, and then have two separate appendices to continue the analysis for each algorithm design we are considering. The almost sure bound for the stability term can be clearly stated as an assumption in the first appendix.}


Recall the definition of the regret
%
\begin{align*}
    R_T &= \bbE \sbr*{\sumtT \spr*{\ell_t \spr*{X_t, A_t} - \ell_t \spr*{X_t, \pi_T^\star \spr*{X_t}}}} + c\,\bbE \sbr*{\sumtT \abs*{O_t}}\\
    &= \bbE \sbr*{\sumtT \sumaK \spr*{q_t \spr*{a \given X_t} - \pi_T^\star \spr*{a \given X_t}} \inp*{X_t, \theta_{t, a}}} + c K\,\bbE \sbr*{\sumtT p_t \spr*{X_t}},
\end{align*}
%
where we used the fact that $\bbE_t \sbr*{\ell_t \spr*{X_t, a}} = \inp*{X_t, \theta_{t, a}}$ for any $a \in \sbr*{K}$ for the first term, and that $\bbE_t \sbr*{\abs*{O_t} \given X_t} = K p_t \spr*{X_t}$ for the second term, which holds because at time $t$, each arm $a$ is observed with probability $p_t \spr*{X_t}$.

We introduce a ghost sample $X_0 \sim \cD$ independent from $\cH_T$. Conditional on $\cH_{t-1}$, both $X_t$ and $X_0$ are \iid from $\cD$, and $p_t$ is $\cH_{t-1}$-measurable, hence, we have
%
\begin{equation*}
    \bbE_t \sbr*{p_t \spr*{X_t}} = \bbE_t \sbr*{p_t \spr*{X_0}}.
\end{equation*}
%
By Lemma~\ref{lem:ghost-sample-neu} and the fact that $\wh{\theta}_{t, a}$ is an unbiased estimator of $\theta_{t, a}$, \ie $\bbE_t \sbr{\wh{\theta}_{t, a}} = \theta_{t, a}$, we can further rewrite the regret as \antoine{there is a bug with the upper bound on the bias term at the end}
%
\begin{align*}
    R_T &= \bbE \sbr*{\sumtT \sumaK \spr*{q_t \spr*{a \given X_0} - \pi_T^\star \spr*{a \given X_0}} \inp*{X_0, \wh{\theta}_{t, a}}} + c K\,\bbE \sbr*{\sumtT p_t \spr*{X_0}}\\
    &= \bbE \sbr*{\sumtT \sumaK \spr*{q_t \spr*{a \given X_0} - \pi_T^\star \spr*{a \given X_0}} \inp*{X_0, \wt{\theta}_{t, a}}} + c K\,\bbE \sbr*{\sumtT p_t \spr*{X_0}}\\
    &\quad+ \bbE \sbr*{\sumtT \sumaK \spr*{q_t \spr*{a \given X_0} - \pi_T^\star \spr*{a \given X_0}} \inp*{X_0, \wh{\theta}_{t, a} - \wt{\theta}_{t, a}}}\\
    &\leq \bbE \sbr*{\sumtT \sumaK \spr*{q_t \spr*{a \given X_0} - \pi_T^\star \spr*{a \given X_0}} \inp*{X_0, \wt{\theta}_{t, a}}} + c K\,\bbE \sbr*{\sumtT p_t \spr*{X_0}}\\
    &\quad+ 2 \sumtT \max_{a \in \sbr*{K}} \abs*{\bbE \sbr*{\inp*{X_0, \wh{\theta}_{t, a} - \wt{\theta}_{t, a}}}}.
\end{align*}
%
For any context $x \in \cX$, we define the auxiliary regret
%
\begin{equation*}
    \wt R_T \spr*{x} \coloneqq \sumtT \sumaK \spr*{q_t \spr*{a \given x} - \pi_T^\star \spr*{a \given x}} \inp*{x, \wt{\theta}_{t, a}} + c K \sumtT p_t \spr*{x}
\end{equation*}
%
and the bias induced by MGR
%
\begin{equation*}
    \biasmgr \coloneqq \sumtT \max_{a \in \sbr*{K}} \abs*{\bbE \sbr*{\inp*{X_0, \wh{\theta}_{t, a} - \wt{\theta}_{t, a}}}}.
\end{equation*}
%
With this notation, we can write the inequality above as
%
\begin{equation*}
    R_T \leq \bbE_{X_0 \sim \cD} \sbr*{\wt R_T \spr*{X_0}} + 2\,\biasmgr.
\end{equation*}
%
We bound the two terms separately. In Lemma~\ref{lem::boundBias} we prove that the bias induced by MGR is upper bounded as
%
\begin{equation*}
    2 \biasmgr \leq \frac{\pi^2}{3}\,.
\end{equation*}
%
Thus, it remains to upper bound $\bbE_{X_0 \sim \cD} \sbr*{\wt R_T \spr*{X_0}}$. Let us fix a context $x \in \cX$ and recall that at time $t$, the distribution $q_t \spr*{\cdot \given x}$ is
%
\begin{equation*}
    q_t \spr*{\cdot \given x} = \argmin_{q \in \Delta_K} \scbr*{\sum_{s=1}^{t-1} \sumaK q \spr*{a} \inp*{x, \wt{\theta}_{s, a}} + \cR_t \spr*{q}},
\end{equation*}
%
where we denoted $\cR_t \spr*{q} = \frac{1}{\eta_t} H_\alpha \spr*{q} + \bar\beta H_{\bar\alpha} \spr*{q}$. By the standard FTRL analysis \citep[Exercise~28.12]{BanditBook} \antoine{copy/prove the lemma in Technical tools appendix and refer to it}, we have
%

\DB{Important point: is the "standard FTRL analysis" mentionned above working just fine if the optimal policy is not unique in some contexts? -- I guess so, because they don't seem to bother in Kuroki-- That is, the FTRL bound holds against any $\pi^\star$?}

\db{Asking because I remember Taira telling me that the uniqueness of the optimal policy was kinda crucial in his decomposition (Ok, actually I recall now more precisely that it was important that it is on a corner of the action set. Here we now that there exists one). However, we can also just keep for now the assumption that there is a gap everywhere (essentially, 
there exists a $\Delta_{\min}>0$), but I suppose that dealing cleanly with multiple optimal actions in some contexts might make the contribution cooler.}

\db{I think there is somewhere a paper on how to cleanly implement/analyze FTRL with multiple optimal actions. Among other things, they do arm-dependent learning rates. Now that I think about this it looks super natural that this is the thing to do to for multiple optimal arms: by having one learning you're kinda aggregating all sub-optimal arms to $1-q_{ta^\star}$.} \DB{Direction for later: might be cool to re-write the FTRL decomposition with arm-dependent learning rate and see what tuning looks natural in each of our variants.}

\db{Given that we use averaging a lot I guess we should write everything with expectations here? Or bound $\bE[\stabt(x)]$ below.}
\begin{align*}
    \wt R_T \spr*{x} &\leq \sumtT \underbrace{\spr*{\sumaK \spr*{q_t \spr*{a \given x} - q_{t+1} \spr*{a \given x}} \inp{x, \wt \theta_{t, a}} - D_{\cR_t} \spr*{q_{t+1} \spr*{\cdot \given x}, q_t \spr*{\cdot \given x}}}}_{= \stabt}\\
    &\quad+ \underbrace{\cR_{T+1} \spr*{\pi_T^\star \spr*{\cdot \given x}} - \cR_1 \spr*{q_1 \spr*{\cdot \given x}} + \sumtT \spr*{\frac{1}{\eta_{t+1}} - \frac{1}{\eta_t}} H_{\alpha} \spr*{q_{t+1} \spr*{\cdot \given x}}}_{= \pen}\\
    &\quad+ c K \sumtT p_t \spr*{x},
\end{align*}
%
where the Bregman divergence of the composite regularizer is given by
%
\begin{equation*}
    D_{\cR_t} \spr*{q, q'} = \cR_t \spr*{q} - \cR_t \spr*{q'} - \inp*{\nabla \cR_t \spr*{q'}, q - q'} = \frac{1}{\eta_t} D_{H_\alpha} \spr*{q, q'} + \bar\beta D_{H_{\bar\alpha}} \spr*{q, q'},
\end{equation*}
%
and the Bregman divergence associated to $H_\alpha$ is
%
\begin{align*}
    D_{H_\alpha} \spr*{q, q'} &= H_\alpha \spr*{q} - H_\alpha \spr*{q'} - \inp*{\nabla H_\alpha \spr*{q'}, q - q'}\\
    &= \frac{1}{\alpha} \sumaK \spr*{q_a - q_a' - q_a^\alpha + \spr*{q_a'}^\alpha + \spr*{\alpha \spr*{q_a'}^{\alpha - 1} - 1} \spr*{q_a - q_a'}}\\
    &= \frac{1}{\alpha} \sumaK \spr*{\spr*{q_a'}^\alpha - q_a^\alpha + \alpha \spr*{q_a'}^{\alpha - 1} \spr*{q_a - q_a'}}.
\end{align*}
\db{use the right entropy}

\paragraph{Controlling the stability term.} Denote $\stab = \sumtT \stabt$. Noting that $D_{H_{\bar\alpha}}$ is non-negative, we have
%
\begin{align*}
    \stabt &\leq \sumaK \spr*{q_t \spr*{a \given x} - q_{t+1} \spr*{a \given x}} \inp{x, \wt \theta_{t, a}} - \frac{1}{\eta_t} D_{H_\alpha} \spr*{q_{t+1} \spr*{\cdot \given x}, q_t \spr*{\cdot \given x}}\\
    &= \frac{1}{\eta_t} \spr*{\sumaK \spr*{q_t \spr*{a \given x} - q_{t+1} \spr*{a \given x}} \eta_t \inp{x, \wt \theta_{t, a}} - D_{H_\alpha} \spr*{q_{t+1} \spr*{\cdot \given x}, q_t \spr*{\cdot \given x}}}.
\end{align*}
%
We need to control the magnitude of $\eta_t \inp{x, \wt \theta_{t, a}}$. Using the definition of $\wt \theta_{t, a}$ and the fact that $\norm{x}_2 \leq 1$, we have \antoine{seems to be the first constraint on the choice of $\eta_t$, we need to set it so we can use Lemma~\ref{lem::ub_ftrl_technical}}
\db{Not on the choice of $\eta_t$ but on the choice of $u_t$. }
To be able to use Lemma~\ref{lem::ub_ftrl_technical}, we need to choose $u_t$ in order to verify that 
%
\begin{equation}\label{eq::as_bound}
    \abs{\eta_t \inp{x, \wt \theta_{t, a}}} \leq \frac{1 - \alpha}{4} \cdot \frac{1}{q_{t\star}(x)^{1 - \alpha}}
\end{equation}
\db{We might naively want to use $u_t\leq \frac{4}{1-\alpha}$, and for the adversarial bound this doesn't seem to cause any harm. However, in Taira's proof, for the stochastic guarantee we need as some point to have $h_t u_t \approx q_{t\star}$, and since $h_t\approx q_{t\star}^\alpha$ the $u_t$ exactly complements it (even though $1-\alpha=\frac{1}{\log(K)}$ is rather small and the difference might appear negligible). But maybe it would be possibly to work the BOBW with $\alpha$ using Jensen or whatever? Investigate.} 
For the moment, forget about MGR and assume that the estimate is of the form $\wt \theta_{t,a} = \Sigma_t^{-1}X_t\ell_t(X_t, a)\ind(a\in O_t)$, where $\Sigma_t$ is any matrix that makes the estimator unbiased/with controlled bias. Then, following e.g the proof of Lemma 2 of \cite{BOBWlinear} we can write that 
\begin{align*}
	\abs{\inp{x, \wt \theta_{t,a}}} &\leq \norm{\Sigma_t^{-1}}_{\text{op}} \\
	& \leq  \frac{1}{\lambda_{\min}(\Sigma_t)} \;, \\
\end{align*}
using that all contexts are in the unit Euclidean ball. Some examples are
\begin{itemize}
	\item[(i)]  $\Sigma_t=p_t(X_t) \Sigma$, which yields $\lambda_{\min}(\Sigma_t) = p_t(X_t)\cdot \lambda_{\min}(\Sigma)$.
	\item[(ii)] $\Sigma_t=\E_{t, X\sim \cD}\left[p_t(X) XX^T\right]$, which gives $\Sigma_t\succeq \lambda_{\min}(\Sigma)\; \underset{x\in \cX}{ \min } \;p_t(x)\cdot  I$ so $\lambda_{\min}(\Sigma_t) \geq \underset{x\in \cX}{ \min }\; p_t(x)  \cdot \lambda_{\min}(\Sigma)$.
	\item[(iii)] MGR estimate of (ii), for which it is known (\cite{BOBWlinear}[Lemma 2]) that $\norm{\Sigma_t^+}_{\text{op}} \leq \frac{(M_t+1)}{2}$.
\end{itemize}

The bound for the MGR estimates encourages to tune $M_t$ at most of order $\frac{1}{\eta_t}$, which is what is done in \cite{BOBWlinear}. This can in turn impose constraints on the tuning of $\eta_t$ (in that case the tuning of $u_t$ will appear in the control of the bias), which gives the extra logs in \cite{BOBWlinear}.  

Apart from the MGR estimate, it seems that the above bounds impose a strong constraint on $p_t$, and so on $u_t$. Indeed, as in \cite{BOBWhardproblems} we use $u_t$ to meet the condition above by choosing it to be $u_t(x)=\eta_t v_t(x)$ for some function $v$. Unfortunately, without proving more properties on $p_t$ it seems that $u_t$ must be uniformly bounded (in both case). In particular, it seems that one must impose 
\begin{equation}\label{eq::cond_ut}\forall x \in \cX, \; u_t(x) = \frac{4}{1-\alpha}\cdot \frac{1}{\lambda_{\min}} \cdot \max_{x\in \cX} q_{t\star}(x)^{1-\alpha} \;. 
\end{equation}

\db{A question we might want to answer is whether using $1/2$ suffices to get some BOBW (at some polylog cost?). This is particularly relevant for future directions to tackle continuous contexts, because have several arms to be optimal/close to optimal would make this constant anyway}


\db{An idea: with the MGR estimate we could reuse the samples to estimate/upper bound this quantity (and the ones we are introducing later). Maybe we can use some sort of smoothness arguments to end up with a reasonable sample size for the approximation. Note that the level of precision could be of order $\sqrt{z_t/\beta_t}$.}


If we can guarantee that \eqref{eq::as_bound} holds, we can then use Lemma~\ref{lem::ub_ftrl_technical} to bound the stability term by 

\begin{equation}\label{eq::stab_bound}
	\stab\leq \frac{4}{1-\alpha} \bE\left[\sum_{t=1}^T  \frac{1}{\beta_t} \bE_t\left[ \sum_{i \neq \wt I_t} q_{t,i}(x)^{2-\alpha} \wt \ell_{t,i}(x)^2 + q_{t\star}(x)^{2-\alpha} \wt \ell_{t,\wt I_t}(x)^2 \right] \right]\;. 
	\end{equation}

To fit in the framework of \cite{BOBWhardproblems} we need to upper bound this term by $\frac{z_t(x)}{\beta_t \gamma'_t(x)}=\sqrt{\frac{z_t(x)}{\beta_t}}$.

Consider first estimate (i). Following Eq.\;(8) from \cite{NeuO20} we can use that, for all $i \in [K]$,  
\begin{align*}
\bE_t\left[\wt \ell_{t,i}(x)^2\right]&\leq\bE_t\left[\frac{\ind(i\in O_t)}{p_t(X_t)^2} x^\top\Sigma^{-1}X_tX_t^\top\Sigma^{-1}x\right] \\
& = \bE_t\left[\frac{\ind(a\in O_t)}{p_t(X_t)^2} x^\top\Sigma^{-1}x \right] \\
& \leq \frac{1}{p_t(X_t)\lambda_{\min}} \;.
%& = \bE_t\left[\frac{\ind(a\in O_t)}{p_t(X_t)^2}\text{Tr}(\Sigma^{-1}x x^\top\Sigma^{-1}X_t X_t^\top)\right] \\
%& \leq \frac{d}{p_t(X_t)} \;.
\end{align*}
Note that compared to \cite{NeuO20} we replace a $d$ by a $\lambda_{\min}^{-1}$ because we cannot take the expectation over $x$ here, and thus apply the trace trick. The change is mild if the distribution of contexts is spanning some origin-centered ball somewhat uniformly, but can be arbitrarily bad if one direction is very rarely sampled.
\db{Encourages us to use the same observation probability for everyone, and get a unique $p_t=\sqrt{\frac{\bE_{X\sim \cD}[z_t(X)]}{\beta_t}}$. Let us elaborate a bit.} \db{Note that we will have to discuss this $\lambda_{\min}$. In the best case this will give us a supplementary $d$ factor compared to e.g. adapting Gergo's alg (my guess).}

Let us introduce 
\[Z_t(x)\coloneqq   \sum_{i \neq \wt I_t} q_{t,i}(x)^{2-\alpha} + q_{t\star}(x)^{2-\alpha}\;. \]
Then, if we fix a shared $p(X_t)=\bar p_t$ for all $t$ and $X_t$, we can obtain from Eq.~\eqref{eq::stab_bound} that 
\[\bE_t[\stabt] \leq \beta_t^{-1} \underbrace{(cK \bar p_t)^{-1}}_{\gamma_t^{-1}} \cdot  \underbrace{\frac{4cKd}{(1-\alpha)\lambda_{\min}}\cdot \bE_{X\sim \cD}[Z_t(X)]}_{z_t} , \] 
if we consider that the ghost sample is independent from the filtration (that is, the filtration is just about things that were seen by the alg). With this writing, the $z_t$ that we fit in the framework of \cite{BOBWhardproblems} is defined as an expectation over contexts, and the SPB matching can then proceed as they describe. \db{Needs to be cleaned/detailed though--TODO.}

We can thus setup 
\[\bar p_t = 2\sqrt{\frac{d \bar Z_t}{cK \lambda_{\min}(\Sigma)(1-\alpha)\beta_t}}+\frac{u_t}{\beta_t}\;. \]

\db{All of this is Ok if we can evaluate it up to a good approximation.} \db{So in summary, with estimate (i) it seems that we can complete the bound on the stability term using the SPB matching definition, and defining $u_t$ as in Eq.~\eqref{eq::cond_ut}. Note that in the case where $p_t$ is uniform then estimates (i) and (ii) match and MGR is not necessary, or just before the interaction to get an approximation of $\Sigma^{-1}$ up to an arbitrary precision.}


On the other hand, using $\Sigma_t$ as in (ii), we get 
\begin{align*}
	\bE_t\left[\wt \ell_{t,i}^2\right]&\leq\bE_t\left[\ind(a\in O_t) x^\top\Sigma_t^{-1}X_tX_t^\top\Sigma_t^{-1}x\right] \\ &  =
	 x^\top\Sigma_t^{-1}\cdot p_t(X_t)X_tX_t^\top\Sigma_t^{-1}x \;.
\end{align*}
We then write that 
\begin{align}\label{eq::stab_context}
	\stab &\leq \frac{4}{1-\alpha} \sum_{t=1}^T \bE\left[\frac{Z_t(x)}{\beta_t p_t(x)} \text{Tr}(p_t(x)xx^\top \Sigma_t^{-1} X_t X_t^\top p_t(X_t)\Sigma_t^{-1})\right] \;.
\end{align}
%




Now to be able to do the same as for Lemma 6 of \cite{NeuO20} we would need to have an  upper bound on $Z_t(x)/p_t(x)$ that is independent of $x$ (but could be a function of e.g. 
$\beta_t$), so the expectation could go in the trace and yield only a $d$ factor. \db{This is what we try to do in the following when using Yuko's approach instead of Taira's. Unless there's something I'm not seeing, I have the impression that Taira's approach is unusable to get to this.}  


\db{Note that in a stochastic setting this tuning kinda makes sense: we want to observe all seemingly sub-optimal arms when we play them (so $p_{t,a}(x))=q_{t,a}(x)$ mimics that), and we want to observe the arm that looks best at a sufficient rate to distinguish it from the others, which is somewhat captured by $q_{t\star}(x)$. Hence I tend to believe that we might be able to make this setting works out. However here forced exploration makes this intuition not as straightforward.} 

\db{Tunnel on the stochastic case: in MAB, the following adaptation of TS should be optimal: 
\begin{itemize}
	\item Say arm $k_t$ has best estimated mean 
	\item Play $k_t$
	\item For observation draw samples $(\wt \theta_{k,t})$ for each arm (i.e. TS), and observe (1) all arms $k \neq k_t$ that beat $\wh \theta_{k_t,t}$ (exploration triggered for sub-optimal arms), and (2) arm $k_t$ if $\wt \theta_{k_t,t}\leq \max_{k\neq k_t} \wh \theta_{k,t}$ (exploration triggered for the optimal arm if it under-performs the best alternative arm's estimate).  
\end{itemize}
The only problem might be linear observation probabilities for the optimal arm if there are two, but maybe adding a threshold is sufficient. If there are two optimal arms we cannot statistically keep the $\log(T)$ rate and it's clear the $T^{2/3}$ must be paid. Last, to make the sampling more cost-aware and interpolate with the full-info regime we might remark that an exploration budget proportional to $1/c$ only costs constant in regret so we can setup an additional layer saying e.g. observe all until you get $1/c$ samples per-arm. \\
Anyway, if we're interested in elaborating I think I can easily write the analysis. End of the tunnel.
}

%We can try to set up (for example) $p_{t,a}(x)=\frac{1}{\sqrt{cK}}q_{t,a}(x)^{2-\alpha}+\frac{\eta_t}{\sqrt{cK}}$ and see how it carries in the analysis. I guess we'd end up with something like 
%\[R_T(x) \lesssim \bE\left[\sum_{t=1}^T \frac{\sqrt{cK}}{\beta_t} + \left(\beta_{t+1}-\beta_t\right) h_t(x) +\frac{\sqrt{cK}}{\beta_t} + \sqrt{cK} Z_t(x) \right]  \;,\]
%which would be, in order, the new versions of the stability, penalty, and exploration term respectively. Note that the only difference with the bound of \cite{BOBWlinear} is the last term: in theirs they only have the forced exploration part. \db{Now the question is: how do we do the learning rate tuning to get what we want?} Note that with the $u_t$ term in the forced exploration we can actually bound the stability by $Z_t(x)\wedge \eta_t$.


\DB{Maybe it would simplify things to use that $h_t(x)$ and $Z_t(x)$ can be quite easily related, remarking that $h_t(x)=\Theta((1-q_{t\star})^{\alpha})$ and $Z_t(x)=\Theta((1-q_{t\star})^{2-\alpha})$. Q: what happens in Taira's analysis if we somewhat replace all $Z_t(x)$ terms as functions of $h_t(x)$? Can we try to come with sthg that looks a bit like Lemma 18 of Kuroki?} \db{Note that this remark is at the core of why our adaptation of Kuroki's approach can work fine.}

\paragraph{Controlling the penalty term.} \db{Previous draft} \antoine{If I understand correctly the second regularizer becomes useful to enable the use of Lemma~\ref{lem::ftrl_smooth_bound}}
\db{The second regularizer controls the ratio of entropies in-between rounds, that is condition (iii) in Thm. 7 of \cite{BOBWhardproblems}.}

\db{Since the learning rate update is defined for SPB matching the only thing to do here is to control this entropy. Seems all arguments from \cite{BOBWhardproblems} should readily apply here with the ghost sample trick.}


\paragraph{Penalty term under $\bar p_t$ algorithm design} If $z_t$ is defined as an expectation over context, then so should be $h_t$, that is 
\[h_t \coloneqq \bE_{x \sim \cD}[h_t(x)] \;, \]
which is what appears in the penalty bound if the learning rate update is independent of contexts.
Since the composite regularizer guarantees that $h_{t+1}(x)\leq 2h_t(x)$ for any context $x$, then the bound holds for $h_t$ too. Hence, this $h_t$ can be used in the learning rate update and condition (iii) of Theorem 7 of \cite{BOBWhardproblems} is satisfied. Main drawback is that this gives us yet another expectation to compute.

\paragraph{What's next} The design discussed above make us fall in the framework of Theorem 7 of \cite{BOBWhardproblems}. 

It seems that we are already satisfying their assumption (i)-(iii) (even if (i) doesn't hold, it can be replaced by "the FTRL regret decomposition holds"), \red{so if we didn't do any mistake so far the bound in $T^{2/3}$ should be fine}. \db{TODO: detail.}

Now it remains to verify if the supplementary conditions needed for the logarithmic bound hold. In Eq. (14) of \cite{BOBWhardproblems}, it is stated that supplementary conditions should be 
\[\sqrt{z_t h_t}\leq \sqrt{\rho_1} (1-q_{t a^\star})  \quad \text{and} \quad u_t h_t \leq \rho_2 (1-q_{ta^\star}) \;. 
\]
What does it mean for us? Well, in both cases the LHS is properly defined, and I assume that for us the condition on the RHS should be taken in expectation over context, to satisfy a somewhat self-bounding property when averaging over contexts. 

If we would like to closely follow Taira's proof I guess we should have that for each individual context $\sqrt{Z_t(x) h_t(x)} \leq 1-q_{t a^\star}(x)$, and so $\bE[\sqrt{Z_t(x) h_t(x)}] \leq \bE[1-q_{t a^\star}(x)]$\footnote{$a^\star$ of course depend on $x$, I'm dropping it from the notation for simplicity}. Unfortunately for us, this is not what we want to verify and Cauchy-Schwarz gives the wrong inequality to get the desired result.

However, we can have some kind of "worst-case" self-bounding property as follows: from (70) and (71) in \cite{BOBWhardproblems} we get that $h_t(x)\lesssim (1-q_{ta^\star}(x))^\alpha$ and $z_t(x)\lesssim (1-q_{ta^\star}(x))^{2-\alpha}$, so we can simply write that 
\begin{align*}
	 z_t h_t & \leq \max_{x \in \cX} z_t(x) \cdot \max_{y \in \cX} h_t(y) \\
	& \leq \max_{x \in \cX}\; (1-q_{ta^\star}(x))^2
\end{align*}
Note that by the same arguments the same type of conditions hold for $u_t h_t$, and actually it's sure we can't do better for this one since we defined $u_t$ precisely by taking this maximum over contexts.

We now show that this "worst-case" self-bounding property is actually sufficient to get a logarithmic bound. Again, the natural inequality is not in the right direction since it is immediate that the regret can be upper bounded by 
\[ R_T \leq 2\sum_{t=1}^T\bE\left[\max_{x \in \cX}(1-q_{ta^\star}(x))\right]\;, \]
but we actually want the reverse inequality. However, we can used that this worst context probability must be matched by some context, occuring with some probability. More precisely, if the context space is discrete for simplicity\footnote{which, again, is not an additional constraint for us since otherwise $\Delta_{\min}$ might be $0$.}, we can get something like 
\[R_T \geq \Delta_{\min} \cdot \inf_{y\in \cX}p_\cD(y)\cdot \bE\left[\max_{x \in \cX}(1-q_{ta^\star}(x))\right] \;,  \]
which is however somewhat super ugly BUT is (I guess) sufficient to claim a log rate when the distributions are stochastic. I assume we just need to replace $\Delta_{\min}$ by $\Delta_{\min} \cdot p_{\cD,\min}$ in Taira's bound. This is not exactly BOBW but it's already a nice result to start with. 

\DB{Question: is it further possible to replace this prob by something proportional to $\lambda_{\min}$ somehow? Might depend on the context distribution, but it'd feel kinda OK imo. Again, this relies on invoking some sort of smoothness of the probabilities in similar directions.}

\subsection{Investigating Kuroki's tuning}


\begin{itemize}
	\item Their Lemma 1 holds for us with $U(x)\to cK U(x)$, bounded by $2cK \bE[\sum_{t=1}^T \gamma_t(X_0)]$, where $\gamma_t(X_0)$ is our observation probability. 
	\item our $u_t/\beta_t$ can be replaced by their $\alpha_t\eta_t$
	\item Their Lemma 8 bounds the stability term. As we described above. The bound becomes $\bE\left[Kd\sum_{t=1}^T \eta_t\right]$ \textcolor{red}{if we set $p_t(x)=\pi_t(a|X_0)$.} \DB{Hmm however it's clear it can't work bc here the exploration would be linear since in this analysis there is no distinction most sampled arm/other arms...looks like Taira's decomposition involving $q_{t\star}$ is important! I'm wondering: can we get something like this with the Shannon entropy too?}
\end{itemize}

\db{Maybe a working idea would be to try to have a learning rate tuning "à la Kuroki" and then trying to verify that all the other terms are bounded as we want. We can try to have a learning rate update that matches the bound of Lemma 5 of Tsuchiya. That is, we know that at the end we want $\bE\left[\sum_{t=1}^T \sqrt{z_t(X_t) h_t(\cdot|X_t)}\right]^{2/3}= \Theta\left(\bE\left[\sum_{t=1}^T \sqrt{ h_t(\cdot|X_t)^\frac{2}{\alpha}}\right]^{2/3}\right)$, so let's approximate $z_t(X_t) \approx h_t(\cdot|X_t)^{\frac{2-\alpha}{\alpha}}$, and find a tuning that would sum as $\bE\left[\sum_{t=1}^T h_t(\cdot|X_t)^{\frac{1}{\alpha}}\right]^{2/3}$ after using similar steps as Lemma 18 of Kuroki.}

Then the idea would be to tune the exploration probabilities in order to simplify the $Z_t(x)$ factor in the stability term, and see what it gives next. 


\paragraph{Starting from the penalty term} We first write that 
\[ \pen =  \bE\left[\sum_{t=1}^T (\beta_{t+1}-\beta_t) H(p_{t+1}(\cdot|X_0))\right] = \bE\left[\sum_{t=1}^T (\beta_{t+1}-\beta_t) H(p_{t+1}(\cdot|X_{t+1}))\right] \;. \]
Then, we see that if we define
\[\beta_{t+1} = \beta_t + \frac{1}{\left(\sum_{s=1}^{t}H(p_s(\cdot|X_s))\right)^{\frac{1}{3}}} \;, \]
we can then obtain that 
\begin{align*}
\pen &= \bE\left[\sum_{t=1}^T \frac{H(p_{t+1}(\cdot|X_{t+1}))}{\left(\sum_{s=1}^{t}H(p_t(\cdot|X_t))\right)^{\frac{1}{3}}}  \right]  \\
& \leq \bE\left[\sum_{t=1}^T \frac{H(p_{t+1}(\cdot|X_{t+1}))}{\left(\sum_{s=1}^{t+1}H(p_s(\cdot|X_s))\right)^{\frac{1}{3}}} \cdot \frac{\left(\sum_{s=1}^{t+1}H(p_s(\cdot|X_s))\right)^{\frac{1}{3}}}{\left(\sum_{s=1}^{t}H(p_s(\cdot|X_s))\right)^{\frac{1}{3}}}  \right] \\
& \leq 2^\frac{1}{3} \cdot \bE\left[\sum_{t=1}^T \frac{H(p_{t+1}(\cdot|X_{t+1}))}{\left(\sum_{s=1}^{t+1}H(p_s(\cdot|X_s))\right)^{\frac{1}{3}}}\right],  \\
\end{align*}
where we upper bounded the second ratio by $2^{1/3}$ using that $\sum_{s=1}^t H(p_s(\cdot|X_s))\geq H(p_1(\cdot|X_1))=h_{\max}$, which is by design of the algorithm, and that $H(p_{t+1}(\cdot|X_{t+1}))\leq h_{\max}$. We further re-index the sum to obtain 
\begin{align*}
\pen &=2^\frac{1}{3} \cdot \left(\bE\left[\sum_{t=1}^{T+1} \frac{H(p_{t}(\cdot|X_{t}))}{\left(\sum_{s=1}^{t}H(p_s(\cdot|X_s))\right)^{\frac{1}{3}}}\right]-1\right). \\
\end{align*}

%and we use the stability property due to the composite regularizer $h_{t+1}(p_{t+1}(\cdot|x)) \leq h_t(p_t(\cdot|x))$, cite Taira's lemma) \db{No be more careful}
We can then use Lemma 12 of \cite{BOBWhardproblems} to upper bound the penalty term by 
\begin{align*}\pen &\leq  \frac{3}{2} \cdot 2^{1/3} \cdot \bE\left[\left(\sum_{t=1}^{T+1}H(p_t(\cdot|X_t))\right)^{\frac{2}{3}} \right]  \\
	& \leq 2 \cdot \left(\bE\left[\sum_{t=1}^{T+1}H(p_t(\cdot|X_t))\right]\right)^{\frac{2}{3}} \\
	& \leq 2 \cdot \left(\bE\left[\sum_{t=1}^{T+1}H(p_t(\cdot|X_0))\right]\right)^{\frac{2}{3}} \\
\end{align*}



which gives us a bound that resembles Lemma 18 of \cite{BOBWlinear}, but with the desired power of $2/3$ instead of $1/2$.

\DB{If everything above works, the composite regularizer might not be necessary: in the proof above we don't need the property $h_{t+1}(x)\leq 2h_t(x)$.}

\db{Note: when it will come to the BOBW bound maybe we will need to have $H^\frac{1}{\alpha}$ in the sum (this is what Taira's proof suggest. See if the above is sufficient, keep that in mind otherwise.}

Then, for the stability term we are interested in lower bounding $\beta_t$. Following \cite{BOBWlinear}, we can use that 
\[\forall t \in [T], \quad \beta_{t+1}- \beta_t \geq \frac{1}{\left(\sum_{s=1}^{t}H(p_s(\cdot|X_s))\right)^{\frac{1}{3}}} \Longrightarrow \frac{1}{\beta_t} \leq \frac{\left(\sum_{s=1}^{t}H(p_s(\cdot|X_s))\right)^{\frac{1}{3}}}{t}\;.  \]
Let see how this gets into the stability term. Let's get back to  
\begin{align*}
	\stab &\leq \frac{4}{1-\alpha} \sum_{t=1}^T \bE\left[\frac{1}{\beta_t} \sum_{a=1}^K \frac{z_{t,a}(x)}{p_{t,a}(x)} \text{Tr}(p_{t,a}(x)xx^\top \Sigma_{t,a}^{-1} X_t X_t^\top p_{t,a}(X_t)\Sigma_{t,a}^{-1})\right] \;,
\end{align*}
To be able to use the trace trick and exhibit a $Kd$ term we need to cancel $z_{t,a}(x)$ with $p_{t,a}(x)$. If we furthermore want to match the stability term to the penalty term it is thus natural to define observation probabilities as 
\[\forall t \in [T], \; \forall \in [K]: \quad p_{t,a}(x) = \frac{z_{t,a}(x)}{\left(\sum_{s=1}^{t}H(p_t(\cdot|X_t))\right)^{\frac{1}{3}}}, \]
which ultimately yields 
\[\stab \leq \frac{4Kd \log(T)}{1-\alpha } \left(\bE\left[\sum_{s=1}^{T}H(p_t(\cdot|X_0))\right]\right)^{\frac{2}{3}} \;. \]
\db{I skipped a couple of steps, but you see that this is essentially the same as above. Again we can adapt constants later to optimize as much as we can. Remark that the $\log(T)$ term is because of the $1/t$. Can be canceled if we add a $t^{\epsilon}$ factor to the observation prob. Check later if we have the slackness to do that.} 

It remains to upper bound the bias term, which is bounded by 
\[\bE\left[\sum_{t=1}^T \gamma_t\right] = cK \bE\left[\sum_{t=1}^T \sum_{a=1}^K p_{t,a}(x)\right] =cK \bE\left[\sum_{t=1}^T \frac{Z_t(x)}{\left(\sum_{s=1}^{t}H(p_t(\cdot|X_t))\right)^{\frac{1}{3}}} \right], \]
which can be bounded exactly the penalty term! This is up to using that $Z_t(x)\lesssim H(p_t(\cdot|x))$, which doesn't seem difficult to show using a couple of Taira's results, and then we use the same trick of replacing $x$ by $X_{t+1}$. 

\db{Do a sanity check of all quantities later. A remark: I didn't handle cleanly the fact that $p_t \leq 1$ and so there must be some clipping in some cases. This is a Todo.} 

\db{Last, I assumed above that the FTRL decomposition holds, which needs Taira's condition on $\eta_t \wt \ell_t$. I assume we really don't want the same tuning as in previous section here, so $u_t$ should be constant/a function of $\beta_t$. However, if we just take $u_t\approx 1$ here we get an additional exploration term that is just $\log(T)T^{1/3}$ at worse here}

\DB{One thing that I don't understand without writing all details is that I don't see where the $\sqrt{\log(K)T}$ bound for the case $c=0$ will appear in this setting. Since our forced exploration term is smaller this should appear somewhere else \dots Investigate this.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
